Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-01-31 src/tex/ms.tex

Since birefringence is a property of spacetime and we are targeting a propagation effect not intrinsic to any source, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
\mi{caveats?}
Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply the product of the individual posteriors,%
\footnote{Had the $\kappa$ prior been nonuniform, then we would need to multiply the individual \emph{likelihoods}, not to the posteriors.}
\begin{equation}
    p(\kappa \mid \{d_i\})\propto \prod_{i}p(\kappa \mid d_i)\,,
    \label{eq:restricted_posterior}
\end{equation}
where $d_i$ is the strain data for the $i$\textsuperscript{th} event.
We use this expression to obtain the primary constraint presented in this work.

@@

With the posterior of $\kappa$ from PE on each event, the posterior of $\mu$ and $\sigma$ can be calculated by
\begin{equation}
    p(\mu,\sigma|\{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i|d_i)}{p(\kappa_i)}p(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
    \label{eq:posterior_of_mu_sigma}
\end{equation}
Note that $p(\kappa)$ is the prior of $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
We chose the prior of $\mu$ and $\sigma$ being uniform distributions.
Equation \ref{eq:posterior_of_mu_sigma} can then be simplified to
\begin{equation}
    p(\mu,\sigma|\{d\})\propto\prod_{i}\int p(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
\end{equation}
To sample the posterior distribution of $\mu$ and $\sigma$, we used a sampling package, flowMC \citep{flowMC}.

We then calculate the generic posterior distribution of $\kappa$ from the posterior of $\mu$ and $\sigma$.
The constraint was calculated by
\begin{equation}
    p(\kappa|\{d\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma|\{d\})\,d\mu\,d\sigma\,.
    \label{eq:generic_posterior}
\end{equation}

>>>>

Since birefringence is a property of spacetime and we are targeting a propagation effect not intrinsic to any source, we should expect $\kappa$ to take the same value for all signals, whether it vanishes or not.
\mi{caveats?}
Under this assumption, the posterior on $\kappa$ inferred collectively from all events with a uniform prior is simply obtained from the product of the individual likelihoods, $p(d_i \mid \kappa) \propto p(\kappa \mid d_i)/p(\kappa)$, such that
\begin{equation}
    p(\kappa \mid \{d_i\})\propto p(\kappa) \prod_{i}\frac{p(\kappa \mid d_i)}{p(\kappa)}\,,
    \label{eq:restricted_posterior}
\end{equation}
where $d_i$ is the strain data for the $i$\textsuperscript{th} event, and $p(\kappa)$ is the prior on $\kappa$; since the prior is uniform, in our case Eq.~\eqref{eq:restricted_posterior} reduces to the product of the posteriors, namely $p(\kappa \mid \{d_i\}) \propto \prod_{i}p(\kappa \mid d_i)$.
We use Eq.~\eqref{eq:restricted_posterior} to obtain the primary constraint presented in this work.

@@

Starting from the posterior on $\kappa$ from each event, the posterior on the hyperparameters $\mu$ and $\sigma$ can be calculated by
\begin{equation}
    p(\mu,\sigma \mid \{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i\mid d_i)}{p(\kappa_i)}p(\kappa_i\mid\mu,\sigma)\,d\kappa_i\,,
    \label{eq:posterior_of_mu_sigma}
\end{equation}
where $p(\kappa)$ is the prior initially applied to $\kappa$ during \ac{PE}, which in our case is a uniform distribution $\mathcal{U}[-1,1]$.
Furhter choosing the hyperpriors on $\mu$ and $\sigma$ to be uniform, Eq.~\eqref{eq:posterior_of_mu_sigma} simplifies to
\begin{equation}
    p(\mu,\sigma\mid\{d\})\propto\prod_{i}\int p(\kappa_i\mid d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
\end{equation}
%
We also calculate the expected population distribution of $\kappa$ marginalized over $\mu$ and $\sigma$.
This is given by
\begin{equation}
            p(\kappa\mid \{d\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma\mid \{d\})\,d\mu\,d\sigma\, , 
    \label{eq:generic_posterior}
\end{equation}
and represents our overall expectation for $\kappa$, given the observed set of individual measurements.
To sample the posterior distribution of $\mu$ and $\sigma$, we use the sampling package \textsc{flowMC} \citep{flowMC}.
